Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-02-10 paper/1_Initial/manuscript.tex
updated manuscript and analysis script

\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, rightrule=.15mm, toprule=.15mm, left=2mm, colback=white, arc=.35mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, opacityback=0]

@@

Interoception is referred to the process of sensing, interpreting and
integrating information pertaining to internal organs, such as the
heart, the lungs or the gut (\citeproc{ref-khalsa2018}{Khalsa et al.,
2018}). While recent research emphasizes a key role of interoception in
a variety of processes (e.g., emotion regulation, decision making) and
of outcomes (physical and psychological well being), the field remains
clouded by concerns about how interoception is assessed.

Various measures of interoception have been developed (see
\textbf{Figure~\ref{fig-measures}}), forming a combination of
``objective'' and ``subjective'' assessments (i.e., physiological tasks
such as the heart beat counting or tracking vs.~questionnaires and
subjective scales involving a metacognitive reflection), ``explicit''
and ``implicit'' paradigms (i.e., directing participants' awareness and
attention to interoceptive processes \emph{vs.} measuring interoception
unbeknownst to them), various interoceptive modalities (e.g.,
cardioception, respiroception, gastroception) and theoretical dimensions
(e.g., accuracy, sensitivity, awareness). While there is no consensus as
to which particular approach provides the most accurate and ``pure''
measure of interoception and interoceptive abilities (assuming it is a
unidimensional construct), it is instead plausible that each measure has
strengths and limitations, and a utility dependent on the context and
goal at hand (\citeproc{ref-jahedi2014}{Jahedi \& Méndez, 2014}).

@@

\includegraphics[width=1\textwidth,height=\textheight]{figures/MeasuringInteroception.jpg}

@@

The use of subjective self-report questionnaires to measure deeply
embodied functions might seem paradoxical. However, recent redefinitions
of interoception emphasize the role of high-level and metacognitive
elaboration of interoceptive information. These redefinitions provide
theoretical grounding to support the idea that some facets of
interoception, including participants' metacognitive beliefs, can be
assessed subjectively. This approach offers useful and interesting
measures (\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The notion that self-reports might not reflect the same processes as
other interoception measures is important to contextualize the apparent
lack of convergence between measures in the field
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}).

A recent systematic review by Desmedt et al.
(\citeproc{ref-desmedt2022measures}{2022}) examined whether various
questionnaires designed to measure ``interoceptive sensibility'' truly
assess the same construct. Notably, Garfinkel et al.
(\citeproc{ref-garfinkel2015knowing}{2015}) defines interoceptive
sensibility as the self-reported tendency to focus on and detect
internal sensations, whereas
(\citeproc{ref-khalsa2018interoception}{\textbf{khalsa2018interoception?}})
defines it more narrowly, excluding the ability to detect these signals.
According to Desmedt et al. (\citeproc{ref-desmedt2022measures}{2022}),
most authors adopt the latter definition. The review found that these
questionnaires measure related but distinct constructs, leading
researchers to commit the `jingle fallacy'---mistakenly treating them as
equivalent measures of interoceptive sensibility.

\textbf{TODO: do a better job}

Few studies have examined the correlation between objective
interoceptive measures, such as the Heartbeat Counting Task (HCT;
Schandry, 1981) and the Heartbeat Detection Task (HDT; Kleckner et al.,
2015). Existing findings report weak correlations for the HCT (Murphy et
al., 2019), no correlations except for the sensitivity variable of the
HCT (Brand et al., 2023), and small correlations for the HDT (Brand et
al., 2023). These results suggest that the IAS, a subjective measure,
does not strongly align with objective interoceptive assessments.

A better understanding of what is being measured with different
questionnaires and dimensions, as well as their potential overlaps with
other constructs (e.g., alexithymia, body awareness), is thus needed to
clarify the role of self-reports in the assessment of interoception.

@@

Although the original validation study suggested a two-factor structure
for the IAS, the authors underline its acceptable but imperfect fit
{[}Murphy et al. (\citeproc{ref-murphy2019}{2019}); p.~127{]}, calling
on further investigation of the scale's factor structure. Notably, the
authors did not define these factors, as no clear explanations were
evident. They suggested that the first factor reflects the perception of
interoceptive signals, while the second pertains to signals that may be
difficult to perceive solely through interoceptive information.

@@

Discussions have also been focused on specific items. For instance,
Murphy et al. (\citeproc{ref-murphy2019}{2019}) notes that some items
might measure direct interoceptive signals such as cardioception, while
others might capture phenomena not perceivable through interoceptive
signals alone (e.g., ``bruising''; p.~119). Lin et al.
(\citeproc{ref-lin2023}{2023}) also highlights their correlation
analysis, showing five locally dependent pairs and three items (touch,
blood sugar, bruise) with exceptionally high difficulty and low
discrimination. Additionally, Campos et al.
(\citeproc{ref-campos2021}{2021}) reported ``tickle'' to be the only
item that reflected more specific factors than the general factor.
Interestingly, Lin et al. (\citeproc{ref-lin2023}{2023}) reported that
all items of the IAS grouped together using a new approach, Exploratory
Graph Analysis {[}EGA; Golino and Epskamp
(\citeproc{ref-golino2017exploratory}{2017}){]}, to assess convergent
and discriminant validity, providing further evidence for
unidimensionality.

The IAS has naturally been compared to other interoception-related
scales, and shows a positive correlations with most facets of the MAIA
(\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}),
except for the Non-Distracting and Not-Worrying subscales
(\citeproc{ref-brand2023}{Brand et al., 2023}). Interestingly, findings
on the correlation between the IAS and the body awareness dimension of
the Body Perception Questionnaire (BPQ-A,
\citeproc{ref-porges1993body}{Porges, 1993}) have been mixed: some
studies report small positive correlations
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), while others find
small negative correlations (\citeproc{ref-lin2023}{Lin et al., 2023})
or no correlation at all (\citeproc{ref-murphy2019}{Murphy et al.,
2019}). Small positive correlations have also been observed with the
``observation'' and ``description'' subscales of the Five Facet
Mindfulness Questionnaire (FFMQ, \citeproc{ref-baer2006using}{Baer et
al., 2006}; \citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), as well as with the
``non-reactivity'' and ``acting with awareness'' subscales
(\citeproc{ref-koike2023}{Koike \& Nomura, 2023}). Additionally, the IAS
has shown a positive correlation with the interoceptive awareness
subscale of the Eating Disorder Inventory (\citeproc{ref-lin2023}{Lin et
al., 2023}) and a negative correlation with the Interoceptive Confusion
Questionnaire (\citeproc{ref-brand2023}{Brand et al., 2023}; ICQ,
\citeproc{ref-brewer2016alexithymia}{Brewer et al., 2016};
\citeproc{ref-murphy2019}{Murphy et al., 2019}). Lastly, small positive
correlations have also been reported with the Interoceptive Attention
Scale (\citeproc{ref-koike2023}{Koike \& Nomura, 2023}; IATS,
\citeproc{ref-lin2023}{Lin et al., 2023}), though studies have also
found no correlation between these measures
(\citeproc{ref-gabriele2022dissociations}{Gabriele et al., 2022}).

While assessing the validity of an interoception scale can be conceived
as theoretically challenging, several measures have been used to assess
convergent validity for the the IAS, including expected negative
associations with alexithymia Murphy et al.
(\citeproc{ref-murphy2019}{2019}), somatic symptoms Lin et al.
(\citeproc{ref-lin2023}{2023}), depressive symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), anxiety
(\citeproc{ref-brand2023}{Brand et al., 2023}), neuroticism
(\citeproc{ref-brand2023}{Brand et al., 2023}) and self-esteem
(\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The current study aims at 1) clarifying the structure of the IAS with a
meta-analytic approach that leverages existing data and contrast the
traditional CFA/SEM factor-based analyses with network-based ones such
as EGA. 2) The second part will provide an overview of the dispositional
correlates of the IAS, providing an overview of the pattern of
associations that is key to better understand the nature, place and role
of interoception questionnaires within a larger context.

@@

Our search focused on studies citing the original IAS validation paper
(\citeproc{ref-murphy2019}{Murphy et al., 2019}), identifying 136 papers
(as of 01/05/2024). To qualify for inclusion, papers needed to (1)
provide accessible data in open-access, (2) employ the IAS as a measure,
and (3) report individual IAS items scores. A total of 10 studies was
included (see \textbf{Table 1}). We also included the data of two
unpublished (but already open-access) studies from the authors and one
from another author. The total N participants was 32,214 participants
(\emph{Mean} = 48.6 years, \emph{SD} = 13.1, 71.6\% Female).

@@

To examine the factor structure of the IAS, a two-step approach was
employed. First, Exploratory Graph Analysis (EGA), was used to estimate
the dimensions via network estimation and community detection, alongside
assessing the stability of dimensions and items using the bootstrapping
techniques (\citeproc{ref-golino2017exploratory}{Golino \& Epskamp,
2017}). The selection of EGA was motivated by its capability to handle
complex, multidimensional data and provide robust dimension estimates. A
novel network psychometrics - Unique variable analysis {[}UVA;
Christensen et al. (\citeproc{ref-christensen2023unique}{2023}){]} -
approach based on the weighted topological overlap will be computed to
evaluate which items have substantial local dependence (\textgreater{}
0.25). Subsequently, exploratory factor analysis (EFA) was employed
followed by confirmatory factor analysis (CFA).

@@

When accounting for all samples, the factor analysis reveals that a
4-factor structure fits best. The exploratory factor analysis revealed
that 4 latent factors (oblimin rotation) accounted for 51.95\% of the
total variance of the original data (MR2 = 19.08\%, MR1 = 17.78\%, MR4 =
9.34\%, MR3 = 5.75\%). Since UVA identified ``tickle'' as the item to be
removed---and it also had the lowest uniqueness value in factor analysis
with a similar loading to ``itch''---``tickle'' was excluded from
subsequent analyses.

Confirmatory factor analysis compared 5 models: a single-factor
solution, a 4-factor solution, a 5-factor solution, a 6-factor solution
and a 7-factor solution. The latter was preferred in most datasets,
including with indices that penalise increased number of parameters
(such as BIC). There was no evidence for higher order factors.

@@

These findings contrast with previous research, which all found that
2-factor model(\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}), 1-factor model
(\citeproc{ref-brand2023}{Brand et al., 2023}) and bifactor model
(\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-lin2023}{Lin et al., 2023}) fits the data best. While this
analysis also revealed an okay fit for the 1-factor model, the 4-factor
model revealed the best fit. The 4-factor structure reveals different
`hubs' of items that are related, not only in this structure analysis,
but also in underlying mechanisms. The `wind-burp-cough-sneeze-vomit'
category, for example, only entails items that are linked to excretion
through the mouth. The other categories are organized similarly. This
organization and structure is useful for further analysis, as the data
can be analyzed and interpreted according to a grouping that is coherent
in result, as well as underlying mechanisms.

There are several items that show redundancy suggesting that adapting
the IAS would be beneficial for validity. Based on the given results, we
suggest removing the tickle, while keeping the itch item
{[}\textbf{todo}: stats?{]}. Other items with slight redundancy were
Hungry and Thirsty, Urinate and Defecate, and Sneeze and Cough.

Interestingly, Lin et al. (\citeproc{ref-lin2023}{2023}) also found that
tickle and itch were redundant, excluding one of them but the reason
being that the character for both words is the same in the Chinese
language. On top of that, they came up with a shortened version of the
IAS, excluding further items, resulting in a 12-item IAS, which aligns
with our findings, suggesting that further items are ambiguous as to
whether they should be removed. In contrast, other findings also found
itch and tickle to be redundant but did not suggest excluding items
(\citeproc{ref-campos2021}{Campos et al., 2021}).

The findings indicate a high proportion of answers at 4 (see Figure 1),
especially when using a 5-step scale. The analogue scale shows a more
dispersed distribution, with some answers indicating the highest 5/5,
which was not the case in the 5-step scale. Therefore, we recommend
using an analog scale for the IAS.

Before this paper, the IAS has not yet been used or analyzed with an
analog scale, rather than a five step scale. Therefore, this study
provides a novel approach to improving the IAS in a simple manner.

@@

There are several limitations to the IAS; There are some redundant
items, the 5-point scale does not provide great variability, and the
structure could be improved. Therefore, improving the IAS, or creating a
new questionnaire investigating interoception could be useful to
achieving reliable and accurate indication of interoceptive awareness.

@@

@@

Correlations will be computed using the correlation package under a
Bayesian framework (ref).

@@

\textbf{todo: compute average correlations }

The EGA components capture groupings of pairs of items, such as ``wind''
and ``burp'', ``cough'' and ``sneeze'', or ``muscle'' and ``pain''.
These groupings were used in the correlational analysis, to observe how
much each group/pair correlates with other factors, such as Alexithymia,
or the MAIA (see Figure 2).

Alexithymia is negatively correlated with all interoceptive groups.
Autistic traits are mostly negatively correlated with IAS measure,
except for patterns and numbers (as an autistic trait), which is
significantly and positively correlated with the ``itch/bruise''
pairing. The BPQ Body Awareness part is positively correlated with all
IAS pairs, whereas the Autonomic Reactivity part is negatively
correlated with all IAS groupings. Conspiracy Beliefs were all
positively correlated with the IAS pairs, however, only Global
Conspiracy with ``hungry/thirsty'', ``urinate/defecate'' and
``cough/sneeze'', as well as personal wellbeing with ``hungry/thirsty'',
and Information Control with ``cough/sneeze'' were significantly
positively correlated. Demographic data is also mostly positively
correlated with the IAS findings, where gender and age more strongly
correlated with ``hungry/thirsty''. In this analysis, lying profile is
not strongly correlated with the IAS; Except for contextuality, which
shows a significant negative correlation with ``itch/bruise''. The MAIA
has a strong positive correlation with most IAS pairings, except for the
``not worrying'' and ``not distracting'' items of the MAIA, which show
less strong, or even negative correlations with all IAS item pairings.
Maladaptiveness had mostly negative correlations with the IAS, with only
a few significant correlations, namely ``psychoticism'', ``negative
affect'', and ``detachment'' with ``muscle/pain'' and
``hungry/thirsty'', as well as ``negative affect'' with ``wind/burp''.
Overall, mood was mostly negatively correlated with the IAS, where
``hungry/thirsty'' had the strongest negative correlation of all mood
measures. Except for ``neuroticism'' and ``honesty-humility'',
personality traits, such as ``openness'' and ``extraversion'' were
positively correlated with the IAS groupings. Shizotypic traits were
mostly negatively correlated with the IAS, with ``hungry/thirsty''
showing the strongest negative correlation between schizotypic and the
IAS. World beliefs were mostly positively correlated with the IAS,
however, only a few were significant: ``hierarchical'', ``enticing'',
and ``alive'' correlates significantly with ``muscle/pain'';
``Understandable'' has a significant positive correlation with
``heart/breathing''; And ``hierarchical'' and ``alive'' has a
significant positive correlation with ``hungry/thirsty''.

>>>>

\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, breakable, opacityback=0, leftrule=.75mm, arc=.35mm, bottomrule=.15mm, toprule=.15mm, left=2mm, colframe=quarto-callout-note-color-frame]

@@

Interoception is referred to the process of sensing, interpreting and
integrating information pertaining to internal organs, such as the
heart, the lungs or the gut (\citeproc{ref-khalsa2018}{Khalsa, Adolphs,
Cameron, Critchley, Davenport, Feinstein, Feusner, Garfinkel, Lane,
Mehling, Meuret, et al., 2018}). While recent research emphasizes a key
role of interoception in a variety of processes (e.g., emotion
regulation, decision making) and of outcomes (physical and psychological
well being), the field remains clouded by concerns about how
interoception is assessed.

Various measures of interoception have been developed (see
Figure~\ref{fig-measures}), forming a combination of ``objective'' and
``subjective'' assessments (i.e., physiological tasks such as the heart
beat counting or tracking vs.~questionnaires and subjective scales
involving a metacognitive reflection), ``explicit'' and ``implicit''
paradigms (i.e., directing participants' awareness and attention to
interoceptive processes \emph{vs.} measuring interoception unbeknownst
to them), various interoceptive modalities (e.g., cardioception,
respiroception, gastroception) and theoretical dimensions (e.g.,
accuracy, sensitivity, awareness). While there is no consensus as to
which particular approach provides the most accurate and ``pure''
measure of interoception and interoceptive abilities (assuming it is a
unidimensional construct), it is instead plausible that each measure has
strengths and limitations, and a utility dependent on the context and
goal at hand (\citeproc{ref-jahedi2014}{Jahedi \& Méndez, 2014}).

@@

\includegraphics[width=6.4in,height=\textheight]{figures/MeasuringInteroception.jpg}

@@

The use of subjective self-report questionnaires to measure deeply
embodied functions might seem paradoxical. However, recent redefinitions
of interoception emphasize the role of high-level and metacognitive
elaboration of interoceptive information. These redefinitions provide
theoretical grounding to support the idea that some facets of
interoception, including participants' metacognitive beliefs, can be
assessed subjectively (\citeproc{ref-khalsa2018interoception}{Khalsa,
Adolphs, Cameron, Critchley, Davenport, Feinstein, Feusner, Garfinkel,
Lane, Mehling, \& others, 2018};
\citeproc{ref-suksasilp2022towards}{Suksasilp \& Garfinkel, 2022}). This
approach offers useful and interesting measures
(\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The notion that self-reports might not reflect the same processes as
other interoception measures is important to contextualize the apparent
lack of convergence between measures in the field
(\citeproc{ref-desmedt2022measures}{Desmedt et al., 2022}). While few
studies have explored the correlation between objective interoceptive
measures, such as the Heartbeat Counting Task (HCT,
\citeproc{ref-schandry1981heart}{Schandry, 1981}) and the Heartbeat
Detection Task (HDT, \citeproc{ref-kleckner2015methodological}{Kleckner
et al., 2015}), and correspondent self-report measures, existing
findings typically show weak or no correlations
(\citeproc{ref-brand2023}{Brand et al., 2023}; e.g., task-based accuracy
vs.~self-reported accuracy, \citeproc{ref-murphy2019}{Murphy et al.,
2019}).

Even among self-report measures low correlations have been observed,
suggesting that these measures, while group in the same umbrella term of
self-report assess different facets of interoception. A recent
systematic review by Desmedt et al.
(\citeproc{ref-desmedt2022measures}{2022}) examined whether various
questionnaires designed to assess ``interoceptive sensibility'' truly
measure the same construct. Notably, Garfinkel et al.
(\citeproc{ref-garfinkel2015knowing}{2015}) defines interoceptive
sensibility as the self-reported tendency to focus on and detect
internal sensations, whereas Khalsa, Adolphs, Cameron, Critchley,
Davenport, Feinstein, Feusner, Garfinkel, Lane, Mehling, and others
(\citeproc{ref-khalsa2018interoception}{2018}) defines it more narrowly,
excluding the ability to detect these signals. According to Desmedt et
al. (\citeproc{ref-desmedt2022measures}{2022}), most authors adopt the
latter definition. The review found that these questionnaires measure
related but distinct constructs, leading researchers to mistakenly
treating them as equivalent measures of interoceptive sensibility. A
better understanding of what is being measured with different
questionnaires and dimensions, as well as their potential overlaps with
other constructs (e.g., alexithymia, body awareness), is thus needed to
clarify the role of self-reports in the assessment of interoception.

@@

Although the original validation study suggested a two-factor structure
for the IAS, the authors underline its acceptable but imperfect fit
(\citeproc{ref-murphy2019}{Murphy et al., 2019, p. 127}), calling on
further investigation of the scale's factor structure. Notably, the
authors did not define these factors, as no clear explanations were
evident. They suggested that the first factor reflects the perception of
interoceptive signals, while the second pertains to signals that may be
difficult to perceive solely through interoceptive information.

\textbf{note to dom}:Morin et al. (\citeproc{ref-morin2016}{2015}) is a
methodological paper on ESEM

@@

Discussions have also been focused on specific items. For instance,
Murphy et al. (\citeproc{ref-murphy2019}{2019}) notes that some items
might measure direct interoceptive signals such as cardioception, while
others might capture phenomena not perceivable through interoceptive
signals alone (e.g., ``bruising''; p.~119). Lin et al.
(\citeproc{ref-lin2023}{2023}) also highlights their correlation
analysis, showing five locally dependent pairs and three items (touch,
blood sugar, bruise) with exceptionally high difficulty and low
discrimination. Additionally, Campos et al.
(\citeproc{ref-campos2021}{2021}) reported ``tickle'' to be the only
item that reflected more specific factors than the general factor.
Interestingly, Lin et al. (\citeproc{ref-lin2023}{2023}) reported that
all items of the IAS grouped together using a new approach, Exploratory
Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{Golino \&
Epskamp, 2017}), to assess convergent and discriminant validity,
providing further evidence for unidimensionality.

The IAS has naturally been compared to other interoception-related
scales, and shows a positive correlations with most facets of the MAIA
(\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}),
except for the Non-Distracting and Not-Worrying subscales
(\citeproc{ref-brand2023}{Brand et al., 2023}). Interestingly, findings
on the correlation between the IAS and the body awareness dimension of
the Body Perception Questionnaire (BPQ-A,
\citeproc{ref-porges1993body}{Porges, 1993}) have been mixed: some
studies report small positive correlations
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), while others find
small negative correlations (\citeproc{ref-lin2023}{Lin et al., 2023})
or no correlation at all (\citeproc{ref-murphy2019}{Murphy et al.,
2019}). Small positive correlations have also been observed with the
``observation'' and ``description'' subscales of the Five Facet
Mindfulness Questionnaire (FFMQ, \citeproc{ref-baer2006using}{Baer et
al., 2006}; \citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023}), as well as with the
``non-reactivity'' and ``acting with awareness'' subscales
(\citeproc{ref-koike2023}{Koike \& Nomura, 2023}). Additionally, the IAS
has shown a positive correlation with the interoceptive awareness
subscale of the Eating Disorder Inventory (\citeproc{ref-lin2023}{Lin et
al., 2023}) and a negative correlation with the Interoceptive Confusion
Questionnaire (\citeproc{ref-brand2023}{Brand et al., 2023}; ICQ,
\citeproc{ref-brewer2016alexithymia}{Brewer et al., 2016};
\citeproc{ref-murphy2019}{Murphy et al., 2019}). Lastly, small positive
correlations have also been reported with the Interoceptive Attention
Scale (\citeproc{ref-koike2023}{Koike \& Nomura, 2023}; IATS,
\citeproc{ref-lin2023}{Lin et al., 2023}), though studies have also
found no correlation between these measures
(\citeproc{ref-gabriele2022dissociations}{Gabriele et al., 2022}).
Interestingly, the IAS and the IATS supposedly measure two different
interoceptive processes (i.e., accuracy and attention, respectively)
which contradict Murphy et al. (\citeproc{ref-murphy2019}{2019})
proposed 2x2 framework.

While assessing the validity of an interoception scale can be conceived
as theoretically challenging, several measures have been used to assess
convergent validity for the the IAS, including expected negative
associations with alexithymia (\citeproc{ref-brand2023}{Brand et al.,
2023}; \citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}), somatic symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), depressive symptoms
(\citeproc{ref-brand2023}{Brand et al., 2023};
\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-lin2023}{Lin et al., 2023}), anxiety
(\citeproc{ref-brand2023}{Brand et al., 2023}), neuroticism
(\citeproc{ref-brand2023}{Brand et al., 2023}) and self-esteem
(\citeproc{ref-murphy2019}{Murphy et al., 2019}).

The current study aims at 1) clarifying the structure of the IAS with a
meta-analytic approach that leverages existing data and contrast the
traditional CFA/SEM factor-based analyses with network-based ones such
as EGA; 2) the second part will provide an overview of the dispositional
correlates of the IAS, clarifying the pattern of associations which is
key to better understand the nature, place and role of interoception
questionnaires within a larger context.

@@

Our search focused on studies citing the original IAS validation paper
(\citeproc{ref-murphy2019}{Murphy et al., 2019}), identifying 136 papers
(as of 01/05/2024). To qualify for inclusion, papers needed to (1)
provide accessible data in open-access, (2) employ the IAS as a measure,
and (3) report individual IAS items scores. A total of 10 studies was
included (see \textbf{Table 1}). We also included the data of two
unpublished (but already open-access) studies from the authors and one
from another researcher. The total N participants was 32,214
participants (\emph{Mean} = 48.6 years, \emph{SD} = 13.1, 71.6\%
Female).

\begin{landscape}

\input{../../analysis/figures/table1.tex}

\end{landscape}

@@

To examine the factor structure of the IAS, a two-step approach was
employed. First, Exploratory Graph Analysis (EGA), was used to estimate
the dimensions via network estimation and community detection, alongside
assessing the stability of dimensions and items using the bootstrapping
techniques (\citeproc{ref-golino2017exploratory}{Golino \& Epskamp,
2017}). The selection of EGA was motivated by its capability to handle
complex, multidimensional data and provide robust dimension estimates. A
novel network psychometrics - Unique variable analysis (UVA,
\citeproc{ref-christensen2023unique}{Christensen et al., 2023}) -
approach based on the weighted topological overlap will be computed to
evaluate which items have substantial local dependence (\textgreater{}
0.25). Subsequently, exploratory factor analysis (EFA) was employed
followed by confirmatory factor analysis (CFA).

@@

When accounting for all samples, the factor analysis reveals that a
4-factor structure fits best. The exploratory factor analysis revealed
that 4 latent factors (oblimin rotation) accounted for 41.58\% of the
total variance of the original data (MR1 = 14.45\%, MR3 = 11.76\%, MR2 =
8.09\%, MR4 = 7.28\%). Since UVA identified ``tickle'' as the item to be
removed---and it also had the lowest uniqueness value in factor analysis
with a similar loading to ``itch''---'' it was excluded from subsequent
analyses.

CFA was computed with the removal of the item ``tickle'' as it was
constantly flagged as redundant. This analysis compared 5 models: a
single-factor solution, a 4-factor solution, a 5-factor solution, a
6-factor solution and a 7-factor solution. The latter was preferred in
most datasets, including with indices that penalise increased number of
parameters (such as BIC). There was no evidence for higher order
factors.

@@

These findings contrast with previous research, which all found that
2-factor model (\citeproc{ref-koike2023}{Koike \& Nomura, 2023};
\citeproc{ref-murphy2019}{Murphy et al., 2019}), 1-factor model
(\citeproc{ref-brand2023}{Brand et al., 2023}) and bifactor model
(\citeproc{ref-campos2021}{Campos et al., 2021};
\citeproc{ref-lin2023}{Lin et al., 2023}) fits the data best. While this
analysis also revealed an okay fit for the 1-factor model, the 4-factor
model was superior. The 4-factor structure reveals different `hubs' of
items that are related, not only in this structure analysis, but also in
underlying mechanisms. The `wind-burp-cough-sneeze-vomit' category, for
example, only entails items that are linked to excretion through the
mouth. The other categories are organized similarly. This organization
and structure is useful for further analysis, as the data can be
analyzed and interpreted according to a grouping that is coherent in
result, as well as underlying mechanisms.

\textbf{note to dom}: what stats do you mean here? the UVA one?

There are several items that show redundancy suggesting that adapting
the IAS would be beneficial for validity. Based on the given results, we
suggest removing the tickle, while keeping the itch item
{[}\textbf{todo}: stats?{]}. Other items with slight redundancy were
``hungry'' and ``thirsty'', ``urinate'' and ``defecate'', and ``sneeze''
and ``cough''.

Interestingly, Lin et al. (\citeproc{ref-lin2023}{2023}) also found that
``tickle'' and ``itch'' were redundant, which led them to excludw one of
them. Although, the reason being that the character for both words is
the same in the Chinese language. On top of that, they came up with a
shortened version of the IAS, excluding further items, resulting in a
12-item IAS, which aligns with our findings, suggesting that further
items are ambiguous as to whether they should be removed. Their 12-item
IAS included the ``hunger'', ``breath'', ``urinate'', ``taste'',
``vomit'', ``cough'', ``temperature'', ``sexual arousa'', ``wind'',
``muscle'', ``pain'', ``itch'' items.\\
In contrast, other findings also found ``tickle'' to be redundant but
did not suggest excluding items (\citeproc{ref-campos2021}{Campos et
al., 2021}).

The findings indicate a high proportion of answers at 4 (see Figure 1),
especially when using a 5-step scale. The analogue scale shows a more
dispersed distribution, with some answers indicating the highest 5/5,
which was not the case in Likert-scales. Therefore, we recommend using
an analog scale for the IAS.

Before this paper, the IAS had not yet been used or analyzed with an
analog scale, rather than a five-step scale. Therefore, this study
provides a novel approach to improving the IAS in a simple manner.

@@

There are several limitations to the IAS. There are some redundant
items, the 5-point scale does not provide great variability, and the
structure could be improved. Therefore, improving the IAS, or creating a
new questionnaire investigating interoception could be useful to
achieving reliable and accurate indication of interoceptive awareness.

@@

\input{../../analysis/figures/table2.tex}

@@

Correlations will be computed using the correlation package under a
Bayesian framework (\citeproc{ref-ben2020effectsize}{Ben-Shachar et al.,
2020}).

@@

\subsection{Average correlations}\label{average-correlations}

The EGA components captured grouping of items such as `wind' and `burp,
'cough' and `sneeze', `muscle' and `pain'. These groupings were used in
correlational analysis to analyse how much each pairing is associated
with other factor such as Alexithymia and with Mood disorders (see
figure 2).

\textbf{note}compute internal consistency

\paragraph{Correlations with body
measures.}\label{correlations-with-body-measures}

Alexithymia was assessed in the samples with the Bermond--Vorst
Alexithymia Questionnaire {[}BVAQ; Vorst and Bermond
(\citeproc{ref-vorst2001validity}{2001}){]} and the Toronto Alexithymia
Scale{[}TAS; Bagby et al. (\citeproc{ref-bagby1994twenty}{1994}){]}.

The BVAQ consists of 5 subscales - fantasising, idenitying, analysing;
verbalising and emotionalising - assessed with 40 items on a 5-point
Likert scale, from `defnitely applies to me' to `in no way applies to
me'. Additionally, the BVAQ reduces these subscales into two high order
factors, an affective component and a cognitive one, with high scores
being indicative of high proneness to alexithymia.

On average, the cognitive component of the BVAQ was weakly and
negatively correlated with all IAS pairs of items with the biggest
correlation being with the Itch/Bruise pair (r = -0.112) and the lowest
correlation beeing with the Muscle/Pain pair (r = -0.244). The affective
component of the BVAQ was positively, but very weak, correlated with all
pairs, with the biggest correlation being with the Itch/Bruise pair (r =
0.107). The only exception was a negative correlation with the
Urinate/Defecate pair (r = -0.036).

The TAS contains 20-items rated on a 5-point forced scale, from
`strongly disagree' to `strongly afree', divided into 3 dimensions -
difficulty identiying feelings, difficulty describing feelings, and
externally thinking. High scores on this scale also reflect higher
alexithymia.

All the three dimensions assessed with the TAS were on average
negatively correlated with all pairs of IAS items. The difficulty
describing feelings had its strongest correlation with Hungry/Thirsty (r
= -0.179) and weakest with the Wind/Burp (r = -0.117). while, the
difficulty describing feelings had its strongest correlation with
Muscle/Pain (r = -0.247) and weakest with Itch/Bruise (r = -0.157).
Lastly, the external thinking dimension was more correlated with the
Cough/Sneeze pair (r = -0.138) and less correlated with the
Hungry/Thirsty (r = -0.018).

The studies within our sample used the Body Perception Questionnaire
short-form (BPQ-SF) and the very-short form (BPQ-VSF) to assess
interoception (\citeproc{ref-cabrera2018assessing}{Cabrera et al.,
2018}). The BPQ-SF comprised of 46 items on a 5-point Likert scale
assessing body awareness (26 items) and autonomic reactivity (21 items).
The BPQ-VSF comprises of 12 items from the body awareness subscale of
the BPQ-SF. In this study, all scores assessing these two dimensions
were grouped together, hence no disitintion is made between awareness
measured with the BPQ-SF and the BPQ-VSF.

\textbf{note to add:} discuss later

In general, all pairs of the IAS were positively,and weakily, associated
with the body awareness subscales, while negative and weakily correlated
with the autonomic reactivity subscale. The strongest correlation
identified between the IAS pairs and the body awereness subscale was
with the\\
Heart/Breathing pair (r = 0.151) whilst the strongest correlation with
the autonomic reactivity was with the Urinate/Defecate pair (r =
-0.235). The weakest correlation between the body awareness and the IAS
was with the Hungry/Thirsty pair (r = 0.055) and between the autonomic
reactivity and the IAS was with the Heart/Breathing pair (r = -0.106).

The MAIA was one of the most commonly used measures of interoception in
our study, with nine samples reporting its use. This 37-item
questionnaire assesses eight state-trait dimensions of interoception:
Noticing, Not-Distracting, Not-Worrying, Attention Regulation, Emotional
Awareness, Self-Regulation, Body Listening, and Trust. Responses are
rated on a scale from 0 (Never) to 5 (Always).

On average, all MAIA dimensions were positively and weakly to moderately
correlated with IAS pairings. Notably, the strongest correlations were
observed between the Noticing dimension and the Heart/Breathing pairing
(r = 0.394), Trusting and Hungry/Thirsty (r = 0.347), and Attention
Regulation and Heart/Breathing (r = 0.334). The Not-Distracting and
Not-Worrying subscales were generally positively correlated with IAS
pairings, with a few exceptions: Not-Distracting showed minimal
correlation with Cough/Sneeze (r = 0.0206) and Heart/Breathing (r =
-0.007), while Not-Worrying had a low correlation with Itch/Bruise (r =
0.031).

The Interoceptive Confusion Questionnaire (ICQ; Brewer et al.
(\citeproc{ref-brewer2016alexithymia}{2016})) was used to assess
individuals' difficulties in interpreting non-affective physiological
states, such as pain and hunger. The ICQ consists of 20 items rated on a
scale from 1 (``Does not describe me'') to 5 (``Describes me very
well''), with higher scores indicating greater interoceptive confusion.
